### 2020-01-06 manuscript/pest-management-science/reviewer-response/reviewer-response.tex
adding to reviewer revisions

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{changepage}% http://ctan.org/pkg/changepage
\usepackage{float}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{graphicx}
\usepackage{ragged2e}
\usepackage{scrextend}
\usepackage{lastpage}
\usepackage{geometry}
 \geometry{
 left = 1.5in,
 right = 1.5in,
 top = 1in
 }
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\rhead{\includegraphics[height= 1cm]{ento1.png}\vspace{1em}}
\chead{\includegraphics[width = 5cm]{agritech1.png}}
\lhead{\includegraphics[width=3cm]{acet1.png}}

\fancyfoot{}
\setlength\headheight{40.8pt}
\RequirePackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{titlesec}
\titleformat{\section}{\normalfont\fontsize{10}{15}\bfseries}{\thesection}{0em}{}
\titlespacing{\section}{}{1.24em}{0.24em}

\usepackage[default]{lato}
\usepackage[T1]{fontenc}

\cfoot{Page \thepage \hspace{1pt} of \pageref{LastPage}}

\begin{document}

\begin{addmargin}[3.8in]{}
Willett, Filgueiras, \newline
Nyrop and Nault \newline
----------------------------- \newline
Denis S. Willett\newline
Department of Entomology \newline
Cornell AgriTech  \newline
Cornell University \newline
15 Castle Creek Drive \newline
Geneva, NY 14456 \newline
deniswillett@cornell.edu \newline
\end{addmargin}
\setlength{\parindent}{0cm}

December 24, 2019

\vspace{1.24em}

Pest Management Science

\vspace{1.24em}

Dear Dr. Torto,

\vspace{0.48em}
\setlength{\parindent}{1.24cm}

Thank you for your consideration of \textit{“Attract and kill: Spinosad containing spheres to control onion maggot (\textit{Delia antiqua})”} for publication in Pest Management Science.  We have revised the manuscript in accordance with reviewer recommendations.  A complete and detailed account of the revisions is included below.  The reviewer comments were very helpful, much appreciated, and have resulted in a better manuscript.  In addition to submitting a revised version of the manuscript incorporating the latest changes, we are also submitting production quality images.

\vspace{2em}

Much appreciated, \newline

\vspace{1em}

Denis S. Willett
\vspace{0.48em}

Camila C. Filgueiras
\vspace{0.48em}

Jan P. Nyrop
\vspace{0.48em}

Brian A. Nault
\vspace{0.48em}

\vspace{1.4em}

\hline

\setlength\parindent{0pt}

\subsection*{Referee 1}

The authors did a great job writing the manuscript.  The introduction, methods, and results were easy to understand.

\begin{quote}
    \textit{Thank you for reading the manuscript and providing constructive feedback.  We appreciate the suggestions and have endeavored to include them in the revised version.  }
\end{quote}

On lines 54 and 55 the authors state field trials were conducted from 2006 to 2008, but they did NOT describe if all experiments were repeated in 2006, 2007, and 2008.  The authors need to describe which experiments were conducted each year, how many times each experiment was replicated, and if the experiments were replicated in space and time.

\begin{quote}
    \textit{Thanks for the opportunity to clarify.  Details on the timing of trials are now included in at the beginning of each sub-section in the Material and Methods section on lines ....  In short, the trials were replicated in space and time.  }
\end{quote}

The authors should add a few sentences describing the field sites (lines 54-56) cropping history (e.g. Where onions grown in previous years?) and the reason why the fields were chosen (e.g. did the fields have a previous history of onion maggot damage?).  If experiments were not replicated in space or time it should be noted in the Material and Methods.

\begin{quote}
    \textit{Thanks for mentioning this opportunity to elaborate.  Details on cropping history and reasoning behind field site choice have now been included on lines ....  Information on how experiments were replicated in space and time are also included (see above).  }
\end{quote}

In the discussion section the authors made no mention of using spinosad seed treatment as a management option in conventional and organic onion fields.  Why?? Spinosad (Regard) seed treatment is currently the primary management method for onion maggot in organic onions, and spinosad is an active ingredient in many conventional seed treatments.   Spinosad is an active ingredient in the FI500 seed treatment, and Regard alone often provides a similar level of protection compared to FI500.  Since FI500 resulted in lower maggot damage compared to all "attract and kill" tactics, why would someone recommend attract and kill with spinosad over spinosad seed treatment?   I strongly recommend the authors discuss the advantages, disadvantages, and compatibility of using spinosad seed treatment and spinosad attract & kill.

The authors should discuss the implications of onion maggot developing resistance to spinosad using the attract and kill tactic.  e.g. Will the attract and kill tactic with spinosad lead to a higher likelihood of onion maggot developing resistance to spinosad?  This tactic targets egg laying adults, it only provides partial control, and it exposes multiple generations per year to spinosad?   These factors seem to favor resistance. This question is especially relevant as the most popular registered insecticide seed treatments contain spinosad and there are few alternatives in the marketplace.

\subsection*{Referee 2}

The methods and results need some significant improvements. I found that methods lacking in details and the analysis not well described. I also noted in the annotated file that the quality of the writing in the intro and discussion is markedly different and will need some work to bring them to the same level.

\begin{quote}
    \textit{Thank you for your comments regarding the manuscript.  We have added extended details to the methods section describing both the field work and analysis.  Additionally, we have adjusted language in the introduction and discussion.  We appreciate your feedback in helping improve the manuscript.  }
\end{quote}

Line 60: Please check and correct.

\begin{quote}
    \textit{Thanks for the opportunity to verify this.  It is correct.  'Abut' (spelled exactly that way) means to share a common border. }
\end{quote}

Line 63: Not sure why its assumed that color and rate are related, explain.

\begin{quote}
    \textit{Interesting question.  There is no assumption that color and rate are related.  The fact that white spheres are attractive to \textit{D. antiqua} is based on previous research in this system.  }
\end{quote}

Line 65: Does this sentence belong to the next paragraph? move it there.

\begin{quote}
    \textit{Excellent suggestion.  This sentence does seem to be hanging there.  Rather than include it in the subsequent paragraph, we have opted to remove it alltogether.  }
\end{quote}

Line 66: be more specific.

\begin{quote}
    \textit{Thanks for the opportunity to elaborate.  We have included more details as to the timing and location of the trials throughout the methods section, specifically lines .... }
\end{quote}

Line 68: climate or weather?

\begin{quote}
    \textit{Interesting question.  We want to use the weather in these trials as a proxy for general exposure to northern climates and have revised the sentence in accordance with your suggestion.  }
\end{quote}

Line 68-69: This was already state above, seems redundant.

\begin{quote}
    \textit{Thanks for mentioning this.  We have removed the redundant statement.  }
\end{quote}

Line 71: More details are needed: how many reps/field, how many fields, how much distance between spheres, etc. Where in the field relative to the crop were spheres? What were they placed on and how high?

\begin{quote}
    \textit{}
\end{quote}

Line 77: Which years was this experiment done? Throughout this section pls change to using SI units.

Line 91: was this based on a preliminary experiment?

Line 99: novel? How was it different from previously used ones?

Line 100: why were these compounds chosen?
\end{document}